Learning action models with minimal observability

### 2019-05-02 doc/aij18/minor-revision/aij18.tex
AIJ18-minor: minor fixes

@@

This great flexibility of \FAMA to different amounts of minimal input knowledge impacts the complexity of the planning tasks. When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, \textcolor{red}{the learning task to be solved can no longer be addressed using current approaches}. This is one of the reasons that justifies our compilation-to-planning scheme in contrast to the most extended SAT-solving scheme used in many existing learning systems. Additionally, our planning-based solving scheme allows us to leverage {\em off-the-shelf} classical planners and benefit from the multiple advances in classical planning research.

>>>>

\textcolor{red}{The lack of observed actions and of a length bound in the input examples is addressed using a compilation-to-planning scheme, which allows a planner to {\em fill the gaps} in the input traces in a way that is consistent with what is observed. Additionally, our planning-based solving scheme allows us to leverage {\em off-the-shelf} classical planners and benefit from the multiple advances in classical planning research.}

%This great flexibility of \FAMA to different amounts of minimal input knowledge impacts the complexity of the planning tasks. When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, \textcolor{red}{the learning task to be solved can no longer be addressed using current approaches}. This is one of the reasons that justifies our compilation-to-planning scheme in contrast to the most extended SAT-solving scheme used in many existing learning systems.

@@

### 2019-05-02 doc/aij18/minor-revision/background.tex
AIJ18-minor: minor fixes

\begin{enumerate}
\item we say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$;  that is $\theta(s_i,\tup{a})=s_{i+1}$. This case clearly states that $\tau_s = s$, meaning that $\forall s_i \in \tau_s$, $s_i$ comprises all the literals of the corresponding state in the trajectory $s$ of the plan $\pi$. Formally, $\forall i, 1 \leq i  < n, |s_i| = |F|$.
\item we say that $\tau_s$ is a partially-observable (\PO) state trajectory if at least one intermediate state of $\tau_s$ is a partial assignment of values to fluents. Formally, $\exists i, 1 \leq i  < n, |s_i| < |F|$. This means that one or more literals are missing in the intermediate $s_i$, all of which may be missing. When all literals are missing, $s_i$ is a \emph{missing} or \emph{empty state} ($s_i = \emptyset$).
\end{enumerate}

@@

\begin{enumerate}
\item when \textbf{all} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, $\tau_s$ is a \emph{non-observable} (\NO) state trajectory. Formally, $\forall i, 1 \leq i < n,  s_i= \emptyset$; i.e., $|s_i| = 0$.
\item when \textbf{none} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, we will refer to $\tau_s$ as a \POstar state trajectory. Formally, $\exists i, 1 \leq i < n, s_i \neq \emptyset$; i.e., $0 < |s_i| < |F|$.
\end{enumerate}

@@

\begin{enumerate}
\item when \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$. In this case, $\tau_a$ contains all the necessary actions to transit every state $s_{i-1}$ to its corresponding successor state $s_{i}$, from $s_0$ to $s_n$. This is the type of input trace accepted by all the existing learning approaches (see section \ref{related_work} for details).
\item when \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence. In this case, at least one of the necessary actions of the plan $\pi$ is missing in $\tau_a$. Formally, $\exists i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.
\item when \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence. Formally, $\forall i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$. That is, $\tau_a = \emptyset$.
\end{enumerate}

>>>>

\begin{enumerate}
\item We say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$;  that is $\theta(s_i,\tup{a})=s_{i+1}$. This case clearly states that $\tau_s = s$, meaning that $\forall s_i \in \tau_s$, $s_i$ comprises all the literals of the corresponding state in the trajectory $s$ of the plan $\pi$. Formally, $\forall i, 1 \leq i  < n, |s_i| = |F|$.
\item We say that $\tau_s$ is a partially-observable (\PO) state trajectory if at least one intermediate state of $\tau_s$ is a partial assignment of values to fluents. Formally, $\exists i, 1 \leq i  < n, |s_i| < |F|$. This means that one or more literals are missing in the intermediate $s_i$, all of which may be missing. When all literals are missing, $s_i$ is a \emph{missing} or \emph{empty state} ($s_i = \emptyset$).
\end{enumerate}

@@

\begin{enumerate}
\item When \textbf{all} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, $\tau_s$ is a \emph{non-observable} (\NO) state trajectory. Formally, $\forall i, 1 \leq i < n,  s_i= \emptyset$; i.e., $|s_i| = 0$.
\item When \textbf{none} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, we will refer to $\tau_s$ as a \POstar state trajectory. Formally, $\exists i, 1 \leq i < n, s_i \neq \emptyset$; i.e., $0 < |s_i| < |F|$.
\end{enumerate}

@@

\begin{enumerate}
\item When \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$. In this case, $\tau_a$ contains all the necessary actions to transit every state $s_{i-1}$ to its corresponding successor state $s_{i}$, from $s_0$ to $s_n$. This is the type of input trace accepted by all the existing learning approaches (see section \ref{related_work} for details).
\item When \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence. In this case, at least one of the necessary actions of the plan $\pi$ is missing in $\tau_a$. Formally, $\exists i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.
\item When \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence. Formally, $\forall i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$. That is, $\tau_a = \emptyset$.
\end{enumerate}

### 2019-05-02 doc/aij18/minor-revision/introduction.tex
AIJ18-minor: minor fixes

In summary, \FAMA is a novel learning approach characterized by:
\begin{itemize}
\item compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces,
\item the plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states,
\item the planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions,
\item a semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model.
\end{itemize}

>>>>

In summary, \FAMA is a novel learning approach characterized by:
\begin{itemize}
\item Compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces.
\item The plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states.
\item The planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions.
\item A semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model.
\end{itemize}

### 2019-05-02 doc/aij18/minor-revision/learning_action_models.tex
AIJ18-minor: minor fixes

When the input plan trace contains observed actions, the extra conditional effects
$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$. \textcolor{red}{As a side note, when the input trace contains all actions (\FO sequence of actions), this conditional effect should be redefined as regular preconditions and effects. Doing so will make it impossible for the planner to execute apply actions outside of the observed ones. Additionally, for traces with \POstar state trajectories, we can force that only one apply action is executed between two observed states by adding the preconditions $\neg action_{applied}$.}\\

>>>>

When the input plan trace contains observed actions, the extra conditional effects
$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$. \textcolor{red}{As a side note, when the input trace contains all actions (\FO sequence of actions), this conditional effect is redefined as regular preconditions and effects. Doing so will make it impossible for the planner to execute apply actions outside of the observed ones. Additionally, for traces with \POstar or \FO state trajectories, we can force that only one apply action is executed between two observed states by adding the preconditions $\neg action_{applied}$.}\\
